Assessing reliability of an online behavioural laterality battery: a pre-registered

### 2020-08-06 online_laterality_reliability.Rmd
Some changes of wording for final check by DB, after incorporating comments from ZW. Now includes mention of Gerrits et al.

Studies of cerebral lateralisation often involve participants completing a series of perceptual tasks under laboratory conditions. This has constrained the number of participants recruited in such studies. Online testing can potentially allow for much larger sample sizes, but limits the amount of experimental control that is feasible. Here we considered whether online testing could give valid and reliable results on four tasks: a rhyme decision visual half-field task, a dichotic listening task, a chimeric faces task, and a finger tapping task to assess manual preference. We recruited 392 participants, oversampling left-handers, who completed the battery twice. Three of the tasks showed evidence of both validity and reliability, insofar as they showed hemispheric advantages in the expected direction and test-retest reliability of at least *r*= .65. The reliability of the rhyme judgement task was less satisfactory (*r*= .63). We also found that left-handers were more likely to depart from typical lateralisation. Contrary to prediction, lateralisation across tasks was not significantly correlated. We conclude that three of the tasks, dichotic listening, chimeric faces and finger tapping, show considerable promise for online evaluation of cerebral lateralisation, but the results do not support the notion of a common underlying cause of lateralisation for different cognitive tasks.

@@

Behavioural laterality tasks have long been used to establish hemispheric dominance for certain cognitive functions. Rather consistently, results from visual half-field and dichotic listening paradigms support the view that the left hemisphere plays the dominant role in language processing (Hugdahl, 2000; Ocklenburg & Güntürkün, 2018). Here we review how each paradigm has been implemented to study language lateralisation and what is known about reliability and validity.

@@

The reliability of the chimeric faces tasks seem promising. Early work by Levy et al (1983) using two vertically presented chimeric faces, found split-half reliability of *r*= .93. The validity of the chimeric faces task has been demonstrated using behavioural and fMRI data, but in a task of recognising facial identity rather than emotion. In Yovel, Tambini, and Brandman’s (2008) study, participants performed a behavioural experiment where they had to recognize centrally presented chimeric faces, which presented different identities in the right and left visual fields. It is encouraging to note that the behavioural measure significantly correlated with the fMRILI calculated from on activation in the fusiform face area (*r*= .49). However, the sample size used here was small: with 17 participants, the 95% confidence interval around the observed correlation of .49 ranged from .012 to .786.

@@

From the above reviewed studies, it appears that handedness is related to laterality across a range of tasks (for further discussion see Carey & Johnstone, 2014). This relationship points towards a general laterality factor which governs lateralisation across tasks and behaviours. This has been argued in several hypotheses (Behrmann & Plaut, 2015; Cook, 1984; Hellige, 1983; Kosslyn, 1987). Common to all hypotheses is the assumption that there would be a correlation between the degree of laterality across complementary tasks. While studies do report significant correlations between measures of lateralisation, these are often small (Badzakova-Trajkov, Haberling, Roberts, & Corballis, 2010). Furthermore, there are a large number of studies which report no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018). However, it has not been possible to clearly unpick whether the null results indicate that there are multiple independent lateral biases across tasks, or whether there is a more prosaic explanation in terms of poor reliability or validity of tasks.

@@

Our final prediction (hypothesis 3) related to the inter-relationships between laterality indices. With online testing, there is the opportunity to test inter-relationships between laterality indices with an adequately powered sample using measures of known reliability. Prior studies have been inconclusive because it has been difficult gather large amounts of data to establish test-retest reliability. For instance, van der Haegen and Brysbaert (2018), found that language laterality did not correlate with face laterality in visual half-field tasks. However, their study was restricted to left-handers, who may be atypical in this regard, and they noted that lack of association was hard to interpret because it may be due to unreliability of the individual measures. In the preregistration for our current study, we predicted that for right-handers, a model where the two language tasks are correlated with each other but not with the chimeric faces (will provide better)model a) would be the best fit to the data, compared to either a model in which laterality indices are correlated between the language and chimeric face tasks (model b), or a base model where each task is independently lateralised (model c). For left-handers, we anticipated that that model c may be the fit to the data, indicating that laterality is dissociated across the different tasks. <!--- ZW: I think you need to explain the rationale for this hypothesis---> Our specfic predictions concerning handedness follow the observations of Woodhead, Bradshaw, Wilson, Thompson, and Bishop (2019) who reported that a single laterality factor was required to explain most the covariance between six language tasks in right-handers. In contrast, two factors were required to explain the covariances in left-handers, indicating that language lateralisation is less consistently lateralised in left-handers.

@@

The battery consisted of eight tasks and were built and administered using the Gorilla Experiment Builder (www.gorilla.sc), a cloud-based tool for collecting data in the behavioural sciences (Anwyl-Irvine, Massonnié, Flitton, Kirkham, & Evershed, 2020). Each task is described below.

@@

The basic demographics questionnaire asked participants to report age, gender, years in education, and whether they were bilingual. We also askedquestions about handedness and footedness. For handedness we asked: which is your dominant hand (i.e. which hand do you prefer to use for tasks such as writing, cutting, and catching a ball)? For footedness we asked: Pwhich foot do you normally use to step up on a ladder/step? For both questions, participants could answer "left", "right", or "no preference".<!---ZW: what options were they given? Left / right / ambidextrous? A reviewer asked me about this recently--->

@@

The Lexical Test for Advanced Learners of English (LexTALE; Lemhöfer & Broersma, 2012) was used to assess participants’ vocabulary knowledge. Participants saw 60 letter strings and had to indicate which word they knew. Of the 60 strings, 40 were actual English words and 20 were non-words. LexTALE scores were corrected for the unequal proportion of words and non-word by averaging the percentages corrected for these two item types (e.g., ((number of words correct/40 x 100) + (number of nonwords correct/20 x 100))/2). Accordingly, the LexTALE scores ranged from 0 to 100.

@@

The rhyme decision task was developed for this project as a novel visual half-field technique which requires participants to decide whether a foveally presented letter string rhymed with the name of an image shown in either the LVF or RVF. With online presentation, a particular concern is potential for eye movements away from fixation. We relied on brief bilateral presentation, which has guaranteed adequate control for eye movements in previous visual half-field experiments with verbal naming (Beaumont, 1982). Studies that directly monitored eye movements in this setting reported failures of fixation on only 0.5% of trials (Geffen, Bradshaw & Nettleton, 1972; but see also Bourne, 2006). The design we used ensured that participants had to start the trial fixating centrally in order to read the target word. Pictures of familiar objects were presented for rhyme matching to the written word, with the goal of ensuring that the participant would have to covertly name the picture, thereby engaging left hemisphere speech processing.

@@

For each task we adopted the following data cleaning procedures. Applying each of these procedures meant that the number of participants left in the final analyses could vary by task. First, for all tasks, we removed trials in which response times were shorter than 200 ms. Response times were log-transformed to reduce skew, as specified in our pre-registration. Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as response times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)). Our differences criterion was set to 1.65 instead of the typical 2.2 because we expected to obtain asymmetric distributions for response time, with most outliers in one tail of the distribution. The Hoaglin and Iglewicz procedure was applied to each participant's data separately. After applying each data trimming procedure, participants left with less than 80% of trials were removed.

For the chimeric faces task, we removed participants if they scored lower than 75% on trials in which an identical emotion was shown in each hemiface, and for the dichotic listening task, we removed those who scored less than 75% correct on trials where the same sound was played to each ear.  These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces and `r nrow(LIs)-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task. An additional 7 cases had to be removed because of a technical error that meant they had less than 80 trials presented.

For the rhyme decision task, we removed participants who scored less than 50% accuracy (chance level) in either visual field. Furthermore, we removed participants who scored less than 75% accuracy on the catch trials (trials in which there was no rhyme). These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on this task. We considered recruiting further participants to achieve our target sample size of 300 participants, but this seemed uneconomic as we could already see that with close to 300 cases for the other tasks that the confidence intervals around the test-retest relibaility estimate were none overalapping with *r*= .65, and we could tell by this point that the test-retest reliability of the rhyme detection task was unsatisfactory; With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r*= `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),2)`, which is smaller than the *r*= 0.65 that we pre-registered as our criterion for a reliable task. We plan to evaluate modified versions of this task in future, but decided to stop data collection at this point and analysed the `r nrow(LIs)` participants with usable data.

@@

The final group of `r nrow(LIs)` participants (`r nrow(LIs[LIs$gender== "Female",])` female) had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female). `r nrow(LIs[LIs$handedness== "Right" & LIs$bilingual== "Yes",])` right-handers were bilingual and `r nrow(LIs[LIs$handedness== "Left" & LIs$bilingual== "Yes",])` left-handers were bilingual.

@@

For the rhyme task, we pre-registered that we would calculate LIs based on log RT data, as accuracy during piloting was at ceiling. To obtain LI_z we ran a *t*-test for each participant where accurate log RTs were the dependent variable and visual field was the independent variable.

@@

To address hypothesis 1 and examine the laterality of each task at the population level, we conducted a series of one-sample *t*-tests for each task using session 1 data.

For the reliability assessment, we checked LIs for normality using a Shapiro-Wilk test. As most LIs were non-normal, we used a series of Spearman’s correlations to examine the correlation of coefficients for session 1 versus session 2. For tasks where the test-retest reliability was above .75 with the existing test length, we reanalysed the data using just the first N trials, reducing N until reliability becomes unsatisfactory (below *r*= .65). This was of interest in indicating the most efficient way to administer these tasks in future, while retaining good reliability. However, the full data (including all trials) was used for all subsequent analyses.

@@

We examined the relationship between LIs on the rhyme decision, dichotic listening, and chimeric faces by modelling covariances using a subset of Structural Equation Modeling (SEM) that does not include any latent variables or directional paths. In this approach, we constrain particular covariance patterns and report ‘best’ model fit according AIC weights (Wagenmakers & Farrell, 2004). Akaike weights are transformations of the AIC values, which can be directly interpreted as conditional probabilities for each model. The best fitting model to the data for right- and left-handers is selected as the final model. This allowed us to compare three models:
-	Model (a), where the two language tasks are correlated with each other but not with the chimeric faces
-	Model (b), where all LIs are correlated, and 
-	Model (c), a base model where the LIs are unrelated. 
This was completed separately for right- and left-handers using both session 1 and session 2 data.

@@

Foot and eye preference characteristics of the sample after exclusions as noted above are shown in Table 1. As participants self-reported eye dominance was always consistent between the Miles and Porta tests, we only report data from the Miles test.

@@

Table 1 <!---New table - lattab to be added in nice format--->
Count of left- and right-handers who report left or right on self-reported footedness, and eyedness (Miles test). 
```{r Tab1,echo=FALSE, comment=FALSE, warning=FALSE}
lattab <- flextable(lattab)
autofit(lattab, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
```
*Note*. RR: right foot, right eye; RL: right foot, left eye; LR: left foot, right eye; LL: left foot, left eye. 'Other' inlcudes those who responded 'Don't know'.

@@

A preliminary impression of the data are shown in the Online Supplemental Materials. The distribution of LIs can be used to quantify departures from symmetry. Recall that for all tasks, we computed LIs in the typical way: (100 x (Right – Left)/(Right + Left)), and that because of crossed pathways from hands, ears and eyes to brain, a positive LI indicates left lateralisation and a negative LI indicates right lateralisation at the brain level. LIs on the rhynme task were calculated using log RTs for correct responses (accuracy was high; *M* = `r round(allmeanpcorr,1)`%, *SD* = `r round(allsdpcorr,1)` %). LIs for the dichotic listening and chieric faces were calculated using the count of correct reponses for sounds/emotions in each ear/visual. LIs on the finger tapping task were calculated using the number of sequences generated for each hand. Table 2 shows these mean LIs, as well as a one-sample Wilcoxon test comparing the mean LI with zero. Wilcoxon tests were chosen as a series of Shapiro-Wilks tests showed departure from normality for a number of tasks. It can be seen that all measures were significantly lateralised in the predicted direction at the population level, with left-hemisphere superiority for rhyme judgement, dichotic listening and finger tapping, and right-hemisphere superiority for chimeric faces.

@@

Table 3 shows data from the LI_z measures, which make it easier to compare different tasks on the same scale. For each task, those with LI_z scores greater than 1.96 are categorised as left-hemisphere dominant, and those with LI_z scores less than -1.96 are categorised as right-hemisphere dominant. If there were no population bias, we would expect around 2.5% of individuals to meet each of these criteria. It can be seen that for dichotic listening and chimeric faces, around 45% of individuals have hemispheric dominance in the predicted direction and around 9% have hemispheric dominance in the opposite direction. The proportions showing significant laterality are lower for the rhyme decision task (around 17% left-hemisphere dominant, and 5% right-hemisphere dominant), and for finger tapping, the proportions are only slightly greater than would be expected by chance, although there is a suggestion of increasing laterality on that task across the two sessions.

```{r make.tab3,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
#Make a table to summarise LI means/distribution stats for 4 tasks this time using LI_z.
#This will gradually get populated as we go through each task
#Cells for each task are filled in course of plots.function chunk
LIztable <- data.frame(matrix(NA,nrow=8,ncol=6))
colnames(LIztable)<-c('Task','N','Mean','SD','L-lateralised %','R-lateralised %')
LIztable[,1]<-c('Rhyme Decision 1','Rhyme Decision 2','Dichotic 1','Dichotic 2','Chimeric Faces 1','Chimeric Faces 2','Finger Tapping 1','Finger Tapping 2')

@@

### Optimising the tasks

The dichotic listening and chimeric faces tasks had test-retest reliability above *r*= .75, and so we investigated whether the tasks could be made more efficient by reducing the number of trials while keeping test-retest reliability above our criterion level of *r*= .65. Hence, we recalculated LIs for the dichotic listening and chimeric faces tasks using 1 to n trials, where n is the total number of trials on each task. The absolute test-retest correlation coefficients are shown in Figure 5. A minimum of 60 trials was required to achieve a test-retest correlation of *r*= .65 using the dichotic listening task. For the chimeric faces task, a minimum of 25 trials is necessary to achieve a test-retest reliability of r= .65. 
<!--- I think this section is useful for us, but maybe not really for others. Is it just diluting the main messages of the paper? --->

![Figure 4. Test-retest correlations for (A) the dichotic listening task and (B) the chimeric faces task when using a variable number of trials to calculate laterality indices. The red dashed line marks the test-retest correlation *r*= .65. The correlations coefficients are shown as absolute values.](all.cor.plot.png)

@@

autofit(ft)
```
*Note*. N is the number of participants. % is the percentage of participants.

@@

In a more formal test we compared three models: (a) where the two language tasks are correlated with eachother but not with the chimeric faces (i.e. a language laterality factor); (b) all LIs are correlations (i.e. a laterality factor); and (c) a base model where the LIs are unrelated (i.e. indepent function). This was done separately for right- and left-handers, using a subset of SEM where we constrain particular covariance patterns and report ‘best’ model fit according to Akaike weights, where a larger weight indicates better fit. The best fitting model to the data for right- and left-handers is selected as the final model.

@@

To examine the relationship between the finger tapping task and a conventional hand preference battery, we computed Spearman's rank correlation between LIs produced on the finger tapping task and the Edinburgh Handedness Inventory. For session 1 data, the relationship between the tasks was moderate, *r*= `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p*< 0.001. For session 2, the correlation was *r*= `r round(speartest2$estimate, 2)`, *95% CI* [`r round(spearCI2$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p*< 0.001. This suggests that the influence of hand preference on the tapping task may increase with practice.

@@

With the advancement of online behaviour research methods, it has become possible to test large samples of participants outside of the lab. In the current study we introduced an online behavioural laterality battery that was administered to a relatively large sample of right- and left-handers. In addition to completing self-report measures of handedness, footedness, and sightedness, participants completed a visual half-field rhyme decision task, a dichotic listening task, a chimeric faces task, and a finger tapping task twice. Thus, we were able to examine the test-retest reliability of each task. We also examined the association between handedness and atypical lateralisation and examined the intercorrelations between LIs on the dichotic listening and chimeric faces tasks. We discuss each of these points in turn.

>>>>

Studies of cerebral lateralisation often involve participants completing a series of perceptual tasks under laboratory conditions. This has constrained the number of participants recruited in such studies. Online testing can potentially allow for much larger sample sizes, but limits the amount of experimental control that is feasible. Here we considered whether online testing could give valid and reliable results on four tasks: a rhyme decision visual half-field task, a dichotic listening task, a chimeric faces task, and a finger tapping task to assess manual preference. We recruited 392 participants, oversampling left-handers, who completed the battery twice. Three of the tasks showed evidence of both validity and reliability, insofar as they showed hemispheric advantages in the expected direction and test-retest reliability of at least *r*= .65. The reliability of the rhyme judgement task was less satisfactory (*r*= .63). We also found that left-handers were more likely to depart from typical lateralisation. Contrary to prediction, lateralisation across tasks was not significantly correlated. We conclude that three of the tasks, dichotic listening, chimeric faces and finger tapping, show considerable promise for online evaluation of cerebral lateralisation, but the results do not support the notion of a common underlying cause of lateralisation for the different cognitive tasks studied here.

@@

Behavioural laterality tasks have long been used to establish hemispheric dominance for certain cognitive functions. Results from visual half-field and dichotic listening paradigms support the view that the left hemisphere plays the dominant role in language processing (Hugdahl, 2000; Ocklenburg & Güntürkün, 2018). Here we review how each paradigm has been implemented to study language lateralisation and what is known about reliability and validity.

@@

The reliability of the chimeric faces tasks seem promising. Early work by Levy et al (1983) using two vertically presented chimeric faces, found split-half reliability of *r*= .93. The validity of the chimeric faces task has been demonstrated using behavioural and fMRI data, but in a task of recognising facial identity rather than emotion. In Yovel, Tambini, and Brandman’s (2008) study, participants performed a behavioural experiment where they had to recognize centrally presented chimeric faces, which presented different identities in the right and left visual fields. It is encouraging to note that the behavioural measure significantly correlated with the fMRI laterality index (LI) calculated from on activation in the fusiform face area (*r*= .49). However, the sample size used here was small: with 17 participants, the 95% confidence interval around the observed correlation of .49 ranged from .012 to .786.

@@

From the above reviewed studies, it appears that handedness is related to laterality across a range of tasks (for further discussion see Carey & Johnstone, 2014). This relationship points towards a general laterality factor which influences lateralisation across tasks and behaviours, as predicted by several theoretical accounts (Behrmann & Plaut, 2015; Cook, 1984; Hellige, 1983; Kosslyn, 1987). In a study of left-handers using fMRI with five tasks, Gerrits, Verhelst, and Vingerhoets (2020) found that the just under half the participants showed either standard laterality (left-lateralised for language and praxis, right-lateralised for face memory, line-bisection and vocal emotion) or a complete reversal of this pattern, providing clear evidence against the idea that these functions are lateralised independently. It is also clear that association is far from perfect,  (Badzakova-Trajkov, Haberling, Roberts, & Corballis, 2010) and some studies have found no correlation between laterality indices (Bryden, Hecaen, & Deagostini, 1983; Rosch, Bishop, & Badcock, 2012; Whitehouse & Bishop, 2009; van der Haegen & Brysbaert, 2018). Our ultimate goal is to unpick whether dissociations between laterality indices are evidence of multiple independent lateral biases across tasks, or whether there is a more prosaic explanation in terms of poor reliability or validity of tasks.

@@

Our final prediction (hypothesis 3) related to the inter-relationships between laterality indices. With online testing, there is the opportunity to test inter-relationships between laterality indices with an adequately powered sample using measures of known reliability. Prior studies have been inconclusive because it has been difficult gather large amounts of data to establish test-retest reliability. For instance, van der Haegen and Brysbaert (2018), found that language laterality did not correlate with face laterality in visual half-field tasks. However, their study was restricted to left-handers, who may be atypical in this regard, and they noted that lack of association was hard to interpret because it may be due to unreliability of the individual measures. In the preregistration for our current study, we predicted that for right-handers, a model where the two language tasks are correlated with each other but not with the chimeric faces (model A) would be the best fit to the data, compared to either a model in which laterality indices are correlated between the language and chimeric face tasks (model B), or a base model where each task is independently lateralised (model C). Given our previous findings using fTCD to study dissociations in language laterality (Woodhead, Bradshaw, Wilson, Thompson, & Bishop, 2019), we further predicted that model fit may differ for right- and left-handers, with more dissociation between tasks in the latter group.

@@

The battery included four measures of background information and four experimental assessments of behaviural laterality that were built and administered using the Gorilla Experiment Builder (www.gorilla.sc), a cloud-based tool for collecting data in the behavioural sciences (Anwyl-Irvine, Massonnié, Flitton, Kirkham, & Evershed, 2020). Each task is described below.

@@

The basic demographics questionnaire asked participants to report age, gender, years in education, and whether they were bilingual. We also asked questions about handedness and footedness. For handedness we asked: which is your dominant hand (i.e. which hand do you prefer to use for tasks such as writing, cutting, and catching a ball)? For footedness we asked: which foot do you normally use to step up on a ladder/step? For both questions, participants could answer "left", "right", or "no preference".

@@

The Lexical Test for Advanced Learners of English (LexTALE; Lemhöfer & Broersma, 2012) was used to screen for normal-range language ability. Participants saw 60 letter strings and had to indicate which words they knew. Of the 60 strings, 40 were actual English words and 20 were non-words. LexTALE scores were corrected for the unequal proportion of words and non-word by averaging the percentages corrected for these two item types (e.g., ((number of words correct/40 x 100) + (number of nonwords correct/20 x 100))/2). Accordingly, the LexTALE scores ranged from 0 to 100.

@@

The rhyme decision task was developed for this project as a novel visual half-field technique designed to engage lateralised language processes. On each trial, participants are asked to decide whether a foveally presented letter string rhymed with the name of an image shown in either the LVF or RVF. With online presentation, a particular concern is potential for eye movements away from fixation. We relied on the fact that participants had to focus initially on the central written word in order to do the task, with pictures subsequently displayed with brief bilateral presentation, which has been shown to give adequate control for eye movements in previous visual half-field experiments with verbal naming (Beaumont, 1982). Studies that directly monitored eye movements in this setting reported failures of fixation on only 0.5% of trials (Geffen, Bradshaw & Nettleton, 1972; but see also Bourne, 2006). Pictures of familiar objects were presented for rhyme matching to the written word, with the goal of ensuring that the participant would have to covertly name the picture, thereby engaging left hemisphere speech processing.

@@

For each task we adopted the following data cleaning procedures. First, for all tasks, we removed trials in which response times were shorter than 200 ms. Response times were log-transformed to reduce skew, as specified in our pre-registration. Outlier reaction times were identified using Hoaglin and Iglewicz’s (1987) procedure, whereby outliers are defined as response times that are 1.65 times the difference between the first and third quantiles (1.65 * (Q3-Q1)) below or above the first and third quantile values (e.g: lower limit = Q1 – 1.65 x (Q3-Q1); upper limit = Q3 + 1.65 x (Q3-Q1)). Our differences criterion was set to 1.65 instead of the typical 2.2 because we expected to obtain asymmetric distributions for response time, with most outliers in one tail of the distribution. The Hoaglin and Iglewicz procedure was applied to each participant's data separately. After applying each data trimming procedure, participants left with less than 80% of trials were removed.

At the next stage, we excluded participants who failed to meet pre-registered criteria for inclusion. Applying each of these procedures meant that the number of participants left in the final analyses could vary by task. For the chimeric faces task, we removed participants if they scored lower than 75% on trials in which an identical emotion was shown in each hemiface, and for the dichotic listening task, we removed those who scored less than 75% correct on trials where the same sound was played to each ear.  These data trimming procedures left data for `r nrow(LIs)-nrow(LIs[LIs$exclude== 1 | LIs$exclude== 12,])` participants on the chimeric faces and `r nrow(LIs)-nrow(LIs[LIs$exclude== 2 | LIs$exclude== 12,])` participants on the dichotic listening task. An additional 7 cases had to be removed because of a technical error that meant they had less than 80 trials presented.

For the rhyme decision task, we removed participants who scored less than 50% accuracy (chance level) in either visual field. Furthermore, we removed participants who scored less than 75% accuracy on the catch trials (trials in which there was no rhyme). These trimming procedures left data for `r nrow(LIs[LIs$exRDT== 0,])` participants on this task. We considered recruiting further participants to achieve our target sample size of 300 participants, but this seemed uneconomic as we could already see that with close to 300 cases, the test-retest reliability of the rhyme detection task was unsatisfactory; With `r nrow(LIs[LIs$exRDT== 0,])` participants, the test-retest reliability was *r* = `r round(cor(LIs[LIs$exRDT== 0,]$Day1.RDT_RT_LI, LIs[LIs$exRDT== 0,]$Day2.RDT_RT_LI, use = "complete.obs"),2)`, which is smaller than the *r* = 0.65 that we pre-registered as our criterion for a reliable task. We plan to evaluate modified versions of this task in future, but decided to stop data collection at this point and so we present here data from those participants with usable data.

@@

The final group consisted of `r nrow(LIs)` participants (`r nrow(LIs[LIs$gender== "Female",])` female) who had complete data on at least one of the behavioural tasks. They had a mean age of `r round(mean(LIs$age, na.rm= TRUE), 1)` years (*SD*= `r round(sd(LIs$age, na.rm= TRUE), 2)` and included `r nrow(LIs[LIs$handedness== "Left",])` left handers (`r nrow(LIs[LIs$handedness== "Left" & LIs$gender== "Female",])` female). `r nrow(LIs[LIs$handedness== "Right" & LIs$bilingual== "Yes",])` right-handers were bilingual and `r nrow(LIs[LIs$handedness== "Left" & LIs$bilingual== "Yes",])` left-handers were bilingual.

@@

For the rhyme task, we pre-registered that we would calculate LIs based on log RT data, as accuracy during piloting was at ceiling. To obtain LI_z we ran a *t*-test comparing accurate log RTs for left- and right-sided stimuli for each person.

@@

To address hypothesis 1 and examine the laterality at the population level, we conducted a series of one-sample *t*-tests to compare mean LIs of the whole sample against zero for each task, using session 1 data.

For the reliability assessment, we first checked LIs for normality using a Shapiro-Wilk test. As most LIs were non-normal, we used a series of Spearman’s correlations to examine the correlation of coefficients for session 1 versus session 2. We also conducted an analysis to determine optimal test length. For tasks where the test-retest reliability was above .75, we reanalysed the data using just the first N trials, reducing N until reliability became unsatisfactory (below *r*= .65). This was of interest in indicating the most efficient way to administer these tasks in future, while retaining good reliability. However, the full data (including all trials) was used for all subsequent analyses.

@@

We examined the relationship between LIs on the rhyme decision, dichotic listening, and chimeric faces by modelling covariances using a subset of Structural Equation Modeling (SEM) that does not include any latent variables or directional paths. In this approach, we constrain particular covariance patterns and report ‘best’ model fit according AIC weights (Wagenmakers & Farrell, 2004). Akaike weights are transformations of the AIC values, which can be directly interpreted as conditional probabilities for each model. The best fitting model to the data for right- and left-handers is selected as the final model. This allowed us to compare three models:  
-	Model A, where the two language tasks are correlated with each other but not with the chimeric faces  
-	Model B, where all LIs are correlated, and   
-	Model C, a base model where the LIs are unrelated.   
This was completed separately for right- and left-handers using both session 1 and session 2 data.

@@

Foot and eye preference characteristics of the sample after exclusions as noted above are shown in Table 1. As participants self-reported eye dominance was always consistent between the Miles and Porta tests, we report data only from the Miles test.

@@

Table 1 <!---New table - lattab to be added in nice format--->
Count of left- and right-handers who report left or right on self-reported footedness, and eyedness (Miles test). 
```{r Tab1,echo=FALSE, comment=FALSE, warning=FALSE}
lattab <- flextable(lattab)
autofit(lattab, add_w = 0.1, add_h = 0.1, part = c("body", "header"))
```
*Note*. RR: right foot, right eye; RL: right foot, left eye; LR: left foot, right eye; LL: left foot, left eye. 'Other' includes those who responded 'Don't know'.

@@

Distributions of the raw data are shown in the Online Supplemental Materials. The distribution of LIs can be used to quantify departures from symmetry. Recall that for all tasks, we computed LIs in the typical way: (100 x (Right – Left)/(Right + Left)), and that because of crossed pathways from hands, ears and eyes to brain, a positive LI indicates left lateralisation and a negative LI indicates right lateralisation at the brain level. LIs on the rhy me task were calculated using log RTs for correct responses (accuracy was high; *M* = `r round(allmeanpcorr,1)`%, *SD* = `r round(allsdpcorr,1)` %). LIs for the dichotic listening and chieric faces were calculated using the count of correct reponses for sounds/emotions in each ear/visual. LIs on the finger tapping task were calculated using the number of sequences generated for each hand. Table 2 shows these mean LIs, as well as a one-sample Wilcoxon test comparing the mean LI with zero. Wilcoxon tests were chosen as a series of Shapiro-Wilks tests showed departure from normality for a number of tasks. It can be seen that all measures were significantly lateralised in the predicted direction at the population level, with left-hemisphere superiority for rhyme judgement, dichotic listening and finger tapping, and right-hemisphere superiority for chimeric faces.

@@

Table 3 shows data from the LI_z measures, which make it easier to compare different tasks on the same scale. For each task, those with LI_z scores greater than 1.96 are categorised as left-hemisphere dominant, and those with LI_z scores less than -1.96 are categorised as right-hemisphere dominant. If there were no population bias, we would expect around 2.5% of individuals to meet each of these criteria. It can be seen that for dichotic listening and chimeric faces, around 45% of individuals have hemispheric dominance in the predicted direction and around 9% have hemispheric dominance in the opposite direction. The proportions showing significant laterality are lower for the rhyme decision task (around 17% left-hemisphere dominant, and 5% right-hemisphere dominant), and for finger tapping, the proportions are only slightly greater than would be expected by chance, although there is a suggestion of increasing laterality on that task across the two sessions. Note, however, that we had oversampled left-handers for this study, and, as is evident from exploratory analyses presented below, finger-tapping was strongly influenced by handedness.
<!---this is also very evident from the scatterplot in Fig 3--->
```{r make.tab3,echo=FALSE,include=TRUE, comment=FALSE, warning=FALSE}
#Make a table to summarise LI means/distribution stats for 4 tasks this time using LI_z.
#This will gradually get populated as we go through each task
#Cells for each task are filled in course of plots.function chunk
LIztable <- data.frame(matrix(NA,nrow=8,ncol=6))
colnames(LIztable)<-c('Task','N','Mean','SD','L-lateralised %','R-lateralised %')
LIztable[,1]<-c('Rhyme Decision 1','Rhyme Decision 2','Dichotic 1','Dichotic 2','Chimeric Faces 1','Chimeric Faces 2','Finger Tapping 1','Finger Tapping 2')

@@

### Optimising the tasks
<!--- DB: I think the key issue is not so much getting reliability of .65, as how many trials could be dropped without any loss of reliability from whatever high level it got to. A bit of rewording to reflect this. Also modified in discussion--->
The dichotic listening and chimeric faces tasks had test-retest reliability above *r* = .75, and so we investigated whether the tasks could be made more efficient by reducing the number of trials while retaining good test-retest reliability. Hence, we recalculated LIs for the dichotic listening and chimeric faces tasks using 1 to n trials, where n is the total number of trials on each task. The absolute test-retest correlation coefficients are shown in Figure 5. Peak reliability was reached in around 85 trials with dichotic listening, and a minimum of 60 trials was required to achieve a test-retest correlation of *r* = .65. For the chimeric faces task, peak reliability was achieve with around 36 trials, and a minimum of 25 trials was necessary to achieve a test-retest reliability of *r* = .65.

![Figure 4. Test-retest correlations for (A) the dichotic listening task and (B) the chimeric faces task when using a variable number of trials to calculate laterality indices. The red dashed line marks the test-retest correlation *r* = .65. The correlations coefficients are shown as absolute values.](all.cor.plot.png)

@@

autofit(ft)
```
*Note*. N is the number of participants. % is the column percentage of participants.

@@

In a more formal test we compared three models: (a) where the two language tasks are correlated with eachother but not with the chimeric faces (i.e. a language laterality factor); (b) all LIs are correlated (i.e. a single laterality factor); and (c) a base model where the LIs are unrelated (i.e. independent functions). This was done separately for right- and left-handers, using a subset of SEM where we constrain particular covariance patterns and report ‘best’ model fit according to Akaike weights, where a larger weight indicates better fit. The best fitting model to the data for right- and left-handers is selected as the final model.

@@

To examine the relationship between the finger tapping task and a conventional hand preference battery, we computed Spearman's rank correlation between LIs produced on the finger tapping task and the Edinburgh Handedness Inventory. For session 1 data, the relationship between the tasks was moderate, *r* = `r round(speartest$estimate, 2)`, *95% CI* [`r round(spearCI$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p* < 0.001. For session 2, the correlation was *r* = `r round(speartest2$estimate, 2)`, *95% CI* [`r round(spearCI2$conf.int[1], 2)`, `r round(spearCI$conf.int[2], 2)`],  *p* < 0.001. This suggests that the influence of hand preference on the tapping task may increase with practice.

@@

With the advancement of online behaviour research methods, it has become possible to test large samples of participants outside the laboratory. In the current study we introduced an online behavioural laterality battery that was administered to a relatively large sample of right- and left-handers. In addition to completing self-report measures of handedness, footedness, and sightedness, participants completed a visual half-field rhyme decision task, a dichotic listening task, a chimeric faces task, and a finger tapping task twice. Thus, we were able to examine the test-retest reliability of each task. We also examined the association between handedness and atypical lateralisation and examined the intercorrelations between LIs on the dichotic listening and chimeric faces tasks. We discuss each of these points in turn.
